---
title: "final_results"
author: "Chris Kardish, Marie Agosta"
date: "December 8, 2016"
output: pdf_document
---
###Results

The results are presented both as they appear in logistic regression ouput (Figure 1) for basic interpretation of the coefficients, but owing to the challenge of interpreting substantive significance from log odds, a table of the same coefficients in exponential form is provided (Figure 3) to show the same results as odds ratios. Confidence intervals are supplied for both the log-odd coefficients (Figure 2) and for odds ratios (Figure 4).  

```{r, include=FALSE}
#setwd('C:/Users/Chris/Documents/GitHub/final_paper')
#setwd('C:/Users/Chris/Documents/R/col_soc_science')
#rm(list = ls()) this command clears your environment
source("cleaning_final.R")

```

```{r, echo=FALSE, message=FALSE, results='asis'}


library(stargazer)

stargazer(M_genY, M_genX, M_boomer, M_silent, type = 'text', 
          dep.var.labels = "Probability of voting in 2012", 
          column.labels = c("Gen Y", "Gen X", "Boomers", "Silent"),
          title = "Figure 1: Log Odds Across Four Generations")
#ci=T
```

```{r, echo=FALSE}

library(knitr)
conf_intervals <- cbind(confint(M_genY), confint(M_genX), confint(M_boomer), 
                        confint(M_silent))
conf_intervals <- round(conf_intervals, digits = 3)
conf_intervals <- as.data.frame(conf_intervals)
colnames(conf_intervals) <- c(1, 2, 3, 4, 5, 6, 7, 8)
conf_intervals$gen_Y <- paste(conf_intervals$`1`, conf_intervals$`2`, sep=",")
conf_intervals$gen_X <- paste(conf_intervals$`3`, conf_intervals$`4`, sep=",")
conf_intervals$boomer <- paste(conf_intervals$`5`, conf_intervals$`6`, sep=",")
conf_intervals$silent <- paste(conf_intervals$`7`, conf_intervals$`8`, sep=",")
varnames <- c("Intercept", "education", "female", "vote_2008", "black", "hispanic", "income")
conf_intervals_kab <- cbind(varnames, conf_intervals$gen_Y, conf_intervals$gen_X,
                        conf_intervals$boomer, conf_intervals$silent)

kable(conf_intervals_kab, format = 'markdown', col.names = c("Coefficients", "Gen Y", 
                "Gen X", "Boomer", "Silent"), 
      caption = "Figure 2: Confidence Intervals of Log Odds")

```

The results in figures 1 and 2 show a considerable degree of consistency across generations. The effect of an additional level of education or income has a similar positive, statistically significant, and substantively significant relationship with voter turnout across generations. The lone exception is education among Generation X, but as seen in Figure 2, the confidence interval does not quite include zero within its range. Still, it is considered statistically insignificant for Generation X. The effect of being a woman is also generally consistent across generations, though not statistically significant, while the effect of past voting behavior (vote_2008) is positive across all generations but varies in terms of the strength of the effect, as will be made clearer below when viewing the exponentiated coefficients. 

The greatest variation between generations can be found in the effects of race, and those effects are not uniformly statistically significant across generations. The effect of being black or hispanic is positive and significant for Millennials, while the effect of being hispanic is not statistically significant or positive for any other generation. The effect of being black is also statistically significant for Baby Boomers and the Silent Generation, but the substantive significance is markedly different: the effect is negative for Baby Boomers but strongly positive among the Silent Generation, even more substantively significant than for Millennials. 

```{r, echo=FALSE, message=FALSE, results='asis'}

stargazer(M_genY, M_genX, M_boomer, M_silent, type = 'text', apply.coef = exp,
         t.auto=F, p.auto=F, report = "vct*", 
         dep.var.labels = "Probability of voting in 2012", 
          column.labels = c("Gen Y", "Gen X", "Boomers", "Silent"),
          title = "Figure 3: Odds Ratios Across Four Generations")

```

```{r, echo=FALSE}

conf_intervals_exp <- cbind(exp(confint(M_genY)), exp(confint(M_genX)), 
                            exp(confint(M_boomer)), exp(confint(M_silent)))
conf_intervals_exp <- round(conf_intervals_exp, digits = 3)
conf_intervals_exp <- as.data.frame(conf_intervals_exp)
colnames(conf_intervals_exp) <- c(1, 2, 3, 4, 5, 6, 7, 8)
conf_intervals_exp$gen_Y <- paste(conf_intervals_exp$`1`, conf_intervals_exp$`2`, sep=",")
conf_intervals_exp$gen_X <- paste(conf_intervals_exp$`3`, conf_intervals_exp$`4`, sep=",")
conf_intervals_exp$boomer <- paste(conf_intervals_exp$`5`, conf_intervals_exp$`6`, sep=",")
conf_intervals_exp$silent <- paste(conf_intervals_exp$`7`, conf_intervals_exp$`8`, sep=",")
conf_intervals_exp_kab <- cbind(varnames, conf_intervals_exp$gen_Y, conf_intervals_exp$gen_X,
                        conf_intervals_exp$boomer, conf_intervals_exp$silent)
kable(conf_intervals_exp_kab, format = 'markdown', col.names = c("Coefficients", "Gen Y", "Gen X", "Boomers", "Silent"), caption = "Figure 4: Confidence Intervals of Odds Ratios")

```

The above findings are simply reinforced when observing the exponentiated coefficients (Figure 3), but their substantive impact is easier to interpret. An additional level of education increases the odds of voting 1.32 times for Millennials, and similarly across all generations. Recall that there are five levels of education, ranging from below high school to graduate-level. An additional level of income increases the odds of voting 1.22 times for Millennials and by a similar factor across all generations. Recall that there are 7 levels of income ranges, from $0-$20,000 per year to $250,000 or above. 

However, it is also true that the odds-ratio of education for Millennials is slightly weaker than that of any other generation, which may reflect diminishing returns of the predictive power of a university education for this generation, which is the best educated of any in American history in terms of bachelor's degree attainment. The gains from a university education grow with the age of the generation, with the strongest odds for the Silent Generation. The effect of income, on the other hand, is stronger for Millennials than any generation other than the Silent Generation, with Generation X and Baby Boomers showing the smallest effects. 

The starkest differences emerge when comparing race. Being black increases the odds of voting relative to the races in the base category (the intercept) -- which include white, Asian, and native American or Pacific islander -- by 2.6. The effect of being black for Baby Boomers decreases the odds of voting relative to other races, while being black and a member of the Silent Generation increases the odds of voting 6.28 times relative to the other races. The effect of being hispanic is statistically significant only for Millennials and is above one only for Millennials as well. With an increase in odds by 1.878 relative races in the base category, the effect for Millennial hispanics is not as strong for blacks. 

Thus, being hispanic is a far stronger predictor of voter turnout for Millennials than for any other generation, and being black is also a stronger predictor for Millennials than for all generations aside from the Silent Generation. The effects of income and education, two well-established predictors of voter turnout, are actually largely consistent across generations. 
