---
title: "final_predictions"
author: "Chris Kardish, Marie Agosta"
date: "December 9, 2016"
output: pdf_document
---
###Predictions using the model
Because of challenges using base R to run and store predictions, this section utilizes the Zelig package and its survey-weighted logit option. It will provide a number of plots showing the predicted probability of a given demographic makeup voting in the 2012 election, based on the model, across a number of different generations but using Millennials as a reference point. The first (below), compares the probability of voting in the 2012 election for a black woman of moderate income ($45,000-$65,000 a year) who did not vote in 2008 at various levels of education. 
```{r, include=FALSE}
#setwd('C:/Users/Chris/Documents/GitHub/final_paper')
#setwd('C:/Users/Chris/Documents/R/col_soc_science')
#rm(list = ls()) this command clears your environment
source("cleaning_final.R")
```

```{r, echo=FALSE, message=FALSE, results='asis', out.width='.49\\linewidth', fig.width=3.5, fig.height=3.5, fig.align='center', fig.show='hold', warning=FALSE}
library(MASS)
library(dplyr)
library(Zelig)
#Steps. 1) run model but using Zelig, which is able to use survey weights
#2) use "setx" function to set the values for the simulation
#3) run the simulation using the parameters you specified (sim)
#4) plot using ci.plot

Z_genY <- zelig(vote_2012 ~ education + female + vote_2008 + black + hispanic + income, model = "logit.survey", weights=~weights, strata=~strata, data = anes_genY)
Z_boomer <- zelig(vote_2012 ~ education + female + vote_2008 + black + hispanic + income, model = "logit.survey", weights=~weights, strata=~strata, 
                  data = anes_boomer)
setZ_genY_black <- setx(Z_genY, education = 1:5, black = 1, hispanic = 0, 
                        income = 3, female = 1, vote_2008 = 0)
simZ_genY_black <- sim(Z_genY, x = setZ_genY_black)
ci.plot(simZ_genY_black, ylim = 0:1, ylab = "Probability of voting", 
        main = "Black Millennial Woman")
setZ_boomer_black <- setx(Z_boomer, education = 1:5, black = 1, hispanic = 0, 
                          income = 3, female = 1, vote_2008 = 0)
simZ_boomer_black <- sim(Z_boomer, x = setZ_boomer_black)
ci.plot(simZ_boomer_black, ylim = 0:1, ylab = "Probability of voting", 
        main = "Black Baby Boomer Woman")
```
The simulation gives a visual sense of the impact of the statistically and substantively significant effect of race for Millennials. A black Millennial woman of moderate income was much more likely to vote in 2012 than a Baby Boomer of similar characteristics across all levels of education. The model cannot explain the reason, however, which would require further research. Given the even stronger statistical and substantive significance of being black for the Silent Generation, it is likely that the predicted probabilities of voting as a black female of moderate education is even higher for this generation. The plot below clearly bears this out.  
```{r, echo=FALSE, message=FALSE, results='asis', out.width='.49\\linewidth', fig.width=3.5, fig.height=3.5, fig.align='center', fig.show='hold', warning=FALSE}
ci.plot(simZ_genY_black, ylim = 0:1, ylab = "Probability of voting", 
        main = "Black Millennial Woman")
Z_silent <- zelig(vote_2012 ~ education + female + vote_2008 + black + hispanic + income, model = "logit.survey", weights=~weights, strata=~strata, 
                  data = anes_silent)
setZ_silent_black <- setx(Z_silent, education = 1:5, black = 1, hispanic = 0, 
                        income = 3, female = 1, vote_2008 = 0)
simZ_silent_black <- sim(Z_silent, x = setZ_silent_black)
ci.plot(simZ_silent_black, ylim = 0:1, ylab = "Probability of voting", 
        main = "Black Silent Generation Woman")
```
We now turn to another demographic. Below are the predicted probabilities of voting in 2012 for a white male with a high school education who did not vote in 2008 across various all levels of income included in the data set.
```{r, echo=FALSE, message=FALSE, results='asis', out.width='.49\\linewidth', fig.width=3.5, fig.height=3.5, fig.align='center', fig.show='hold', warning=FALSE}

setZ_genY_white <- setx(Z_genY, income = 1:7, education = 2, black = 0, 
                        hispanic = 0, female =0, vote_2008 = 0)
simZ_genY_white <- sim(Z_genY, x = setZ_genY_white)
ci.plot(simZ_genY_white, ylim = 0:1, ylab = "Probability of voting", 
        main = "White Millennial Man")
#Now, Baby Boomer
setZ_boomer_white <- setx(Z_boomer, income = 1:7, education = 2, black = 0, 
                          hispanic = 0, female =0, vote_2008 = 0)
simZ_boomer_white <- sim(Z_boomer, x = setZ_boomer_white)
ci.plot(simZ_boomer_white, ylim = 0:1, ylab = "Probability of voting", 
        main = "White Baby Boomer Man")
```
In the case of white men with only a high-school education the model predicts essentially the same starting point of around 37 percent for Millennials and Baby Boomers at the lowest income range of 0-$25,000. But the Millennial in this case quickly surpasses the Baby Boomer in probability of voting in 2012, even if not by a substantial percentage, ultimately topping out at about 60 percent at the highest level of income. This likely reflects the heightened importance of income for the Millennials in terms of predicting 2012 voter turnout. Still, the difference is only slight, and the differences between Millennials and the other two generations for this demographic are similarly quite minor. They are shown below.
```{r, echo=FALSE, message=FALSE, results='asis', out.width='.49\\linewidth', fig.width=3.5, fig.height=3.5, fig.align='center', fig.show='hold', warning=FALSE}
ci.plot(simZ_genY_white, ylim = 0:1, ylab = "Probability of voting", 
        main = "White Millennial Man")
Z_genX <- zelig(vote_2012 ~ education + female + vote_2008 + black + hispanic + income, model = "logit.survey", weights=~weights, strata=~strata, 
                  data = anes_genX)

setZ_genX_white <- setx(Z_genX, income = 1:7, education = 2, black = 0, 
                        hispanic = 0, female =0, vote_2008 = 0)
simZ_genX_white <- sim(Z_genX, x = setZ_genX_white)
ci.plot(simZ_genY_white, ylim = 0:1, ylab = "Probability of voting", 
        main = "White Gen X Man")
```

```{r, echo=FALSE, message=FALSE, results='asis', out.width='.49\\linewidth', fig.width=3.5, fig.height=3.5, fig.align='center', fig.show='hold', warning=FALSE}
ci.plot(simZ_genY_white, ylim = 0:1, ylab = "Probability of voting", 
        main = "White Millennial Man")
setZ_silent_white <- setx(Z_silent, income = 1:7, education = 2, black = 0, 
                        hispanic = 0, female =0, vote_2008 = 0)
simZ_silent_white <- sim(Z_silent, x = setZ_silent_white)
ci.plot(simZ_genY_white, ylim = 0:1, ylab = "Probability of voting", 
        main = "White Silent Generation Man")
```
The other variables that shows significant variation between generations in terms of substantive significance is past voting history. As was noted in the results section, the odds ratio for having voted in 2008 is positively statistically and substantively significant across all generations, but its intensity increases with the age of the generation, with the strongest effects observed for the Silent Generation. Those differences are clearly observed in the predicted-probability plots below, which continue use the same demographic as above but now assume that this person voted in 2008. 
```{r, echo=FALSE, message=FALSE, results='asis', out.width='.49\\linewidth', fig.width=3.5, fig.height=3.5, fig.align='center', fig.show='hold', warning=FALSE}

setZ_genY_voted <- setx(Z_genY, income = 1:7, education = 2, black = 0, 
                        hispanic = 0, female =0, vote_2008 = 1)
simZ_genY_voted <- sim(Z_genY, x = setZ_genY_voted)
ci.plot(simZ_genY_voted, ylim = 0:1, ylab = "Probability of voting", 
        main = "White Millennial man who voted in '08")

setZ_genX_voted <- setx(Z_genX, income = 1:7, education = 2, black = 0, 
                        hispanic = 0, female =0, vote_2008 = 1)
simZ_genX_voted <- sim(Z_genX, x = setZ_genX_voted)
ci.plot(simZ_genX_voted, ylim = 0:1, ylab = "Probability of voting", 
        main = "White Gen X man who voted in '08")
```
The difference in probability between Millennials and Generation X is clear, and it is due to the influence of past voting decisions in determining future turnout. Below we instead compare Generation X with Baby Boomers and then Baby Boomers with the Silent Generation to show how the predicted probability of voting in 2012 intensifies with age when assuming that the respondent voted in 2008. 
```{r, echo=FALSE, message=FALSE, results='asis', out.width='.49\\linewidth', fig.width=3.5, fig.height=3.5, fig.align='center', fig.show='hold', warning=FALSE}
ci.plot(simZ_genX_voted, ylim = 0:1, ylab = "Probability of voting", 
        main = "White Gen X man who voted in '08")
setZ_boomer_voted <- setx(Z_boomer, income = 1:7, education = 2, black = 0, 
                          hispanic = 0, female =0, vote_2008 = 1)
simZ_boomer_voted <- sim(Z_boomer, x = setZ_boomer_voted)
ci.plot(simZ_boomer_voted, ylim = 0:1, ylab = "Probability of voting", 
        main = "White Baby Boomer man who voted in '08")
```

```{r, echo=FALSE, message=FALSE, results='asis', out.width='.49\\linewidth', fig.width=3.5, fig.height=3.5, fig.align='center', fig.show='hold', warning=FALSE}
ci.plot(simZ_boomer_voted, ylim = 0:1, ylab = "Probability of voting", 
        main = "White Baby Boomer man who voted in '08")

setZ_silent_voted <- setx(Z_silent, income = 1:7, education = 2, black = 0, 
                          hispanic = 0, female =0, vote_2008 = 1)
simZ_silent_voted <- sim(Z_silent, x = setZ_silent_voted)
ci.plot(simZ_silent_voted, ylim = 0:1, ylab = "Probability of voting", 
        main = "White Silent Generation man who voted in '08")

```

