---
title: "Webpage"
author: "me"
output: 
  flexdashboard::flex_dashboard:
    orientation: rows
    social: menu
    source_code: embed
runtime: shiny

---
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
#setwd('C:/Users/Chris/Documents/GitHub/final_paper')
source("cleaning_final.R")
#rm(list = ls()) 
library("flexdashboard")
```

Inputs {.sidebar}
-----------------------------------------------------------------------

```{r}
 selectInput("clusterNum", label = ("Basic_Information"), 
    choices = list("1" = 1, "2" = 2, "3" = 3, "4" = 4, "5" = 5), 
    selected = 1)
```

## Research Question 

 **In the US, do traditional indicators of individual voter turnout hold the same predictive power across generations?** 

Traditional indicators of voter turnout: income, age, occupation, education, past voting decisions

Approximately **69.2 million Millenials** are of voting age and are eligible to vote, making them the largest share of eligible voters. Yet...

- In the 2008 presidential election, 50% voted 
- In the 2012 presidential election, 46% voted 
- In the 2016 presidential election, had only millennials voted, Clinton would've won the election in a landslide - 473 electoral votes to Trump's 32.

```{r}
 selectInput("clusterNum", label = h3("Cluster number"), 
    choices = list("1" = 1, "2" = 2, "3" = 3, "4" = 4, "5" = 5), 
    selected = 1)
```


Row
-----------------------------------------------------------------------

### Odds-Ratios

    
```{r echo = FALSE, results='asis', message=FALSE,}
#setwd('C:/Users/Chris/Documents/GitHub/Final_Presentation')
library(stargazer)
stargazer(M_genY, M_genX, M_boomer, M_silent, apply.coef = exp,
         t.auto=F, p.auto=F, report = "vct*", font.size = 'tiny',
          type = "html",  dep.var.labels = "Probability of voting in 2012", 
          column.labels = c("Gen Y", "Gen X", "Boomers", "Silent"),
          title = "Odds Ratios Across Four Generations")
```
```
    
Row {.tabset}
-----------------------------------------------------------------------

### Black Women, Moderate Income 

```{r, echo=FALSE, include=FALSE, message=FALSE}
library(Zelig)
Z_genY <- zelig(vote_2012 ~ education + female + vote_2008 + black + hispanic + income, model = "logit.survey", weights=~weights, strata=~strata, data = anes_genY)
Z_boomer <- zelig(vote_2012 ~ education + female + vote_2008 + black + hispanic + income, model = "logit.survey", weights=~weights, strata=~strata, 
                  data = anes_boomer)
Z_silent <- zelig(vote_2012 ~ education + female + vote_2008 + black + hispanic + income, model = "logit.survey", weights=~weights, strata=~strata, 
                  data = anes_silent)
Z_genX <- zelig(vote_2012 ~ education + female + vote_2008 + black + hispanic + income, model = "logit.survey", weights=~weights, strata=~strata, 
                  data = anes_genX)
```
```{r, message=FALSE, echo = FALSE, fig.align='center', fig.show='hold', fig.width=4, fig.height=5}
setZ_genY_black <- setx(Z_genY, education = 1:5, black = 1, hispanic = 0, 
                        income = 3, female = 1, vote_2008 = 0)
simZ_genY_black <- sim(Z_genY, x = setZ_genY_black)
ci.plot(simZ_genY_black, ylim = 0:1, ylab = "Probability of voting", 
        main = "Black Millennial Woman")
setZ_boomer_black <- setx(Z_boomer, education = 1:5, black = 1, hispanic = 0, 
                          income = 3, female = 1, vote_2008 = 0)
simZ_boomer_black <- sim(Z_boomer, x = setZ_boomer_black)
ci.plot(simZ_boomer_black, ylim = 0:1, ylab = "Probability of voting", 
        main = "Black Baby Boomer Woman")
```

### Other plot 

```{r, message=FALSE, echo = FALSE, fig.align='center', fig.show='hold', fig.width=3.5, fig.height=5}
ci.plot(simZ_genY_black, ylim = 0:1, ylab = "Probability of voting", 
        main = "Black Millennial Woman")
setZ_silent_black <- setx(Z_silent, education = 1:5, black = 1, hispanic = 0, 
                        income = 3, female = 1, vote_2008 = 0)
simZ_silent_black <- sim(Z_silent, x = setZ_silent_black)
ci.plot(simZ_silent_black, ylim = 0:1, ylab = "Probability of voting", 
        main = "Black Silent Generation Woman")
```

### different

```{r, echo=FALSE, include=FALSE, message=FALSE}
library(Zelig)
Z_genY <- zelig(vote_2012 ~ education + female + vote_2008 + black + hispanic + income, model = "logit.survey", weights=~weights, strata=~strata, data = anes_genY)
Z_boomer <- zelig(vote_2012 ~ education + female + vote_2008 + black + hispanic + income, model = "logit.survey", weights=~weights, strata=~strata, 
                  data = anes_boomer)
Z_silent <- zelig(vote_2012 ~ education + female + vote_2008 + black + hispanic + income, model = "logit.survey", weights=~weights, strata=~strata, 
                  data = anes_silent)
Z_genX <- zelig(vote_2012 ~ education + female + vote_2008 + black + hispanic + income, model = "logit.survey", weights=~weights, strata=~strata, 
                  data = anes_genX)
```
```{r, message=FALSE, echo = FALSE, fig.align='center', fig.show='hold', fig.width=4, fig.height=5}
setZ_genY_black <- setx(Z_genY, education = 1:5, black = 1, hispanic = 0, 
                        income = 3, female = 1, vote_2008 = 0)
simZ_genY_black <- sim(Z_genY, x = setZ_genY_black)
ci.plot(simZ_genY_black, ylim = 0:1, ylab = "Probability of voting", 
        main = "Black Millennial Woman")
setZ_boomer_black <- setx(Z_boomer, education = 1:5, black = 1, hispanic = 0, 
                          income = 3, female = 1, vote_2008 = 0)
simZ_boomer_black <- sim(Z_boomer, x = setZ_boomer_black)
ci.plot(simZ_boomer_black, ylim = 0:1, ylab = "Probability of voting", 
        main = "Black Baby Boomer Woman")
```

### Other plot  

```{r, message=FALSE, echo = FALSE, fig.align='center', fig.show='hold', fig.width=3.5, fig.height=5}
ci.plot(simZ_genY_black, ylim = 0:1, ylab = "Probability of voting", 
        main = "Black Millennial Woman")
setZ_silent_black <- setx(Z_silent, education = 1:5, black = 1, hispanic = 0, 
                        income = 3, female = 1, vote_2008 = 0)
simZ_silent_black <- sim(Z_silent, x = setZ_silent_black)
ci.plot(simZ_silent_black, ylim = 0:1, ylab = "Probability of voting", 
        main = "Black Silent Generation Woman")
```

###  white, high-school-educated man across levels of income, assuming he didn't vote in 2008. 

```{r, message=FALSE, echo = FALSE, fig.align='center', fig.show='hold', fig.width=3.5, fig.height=5}
setZ_genY_white <- setx(Z_genY, income = 1:7, education = 2, black = 0, 
                        hispanic = 0, female =0, vote_2008 = 1)
simZ_genY_white <- sim(Z_genY, x = setZ_genY_white)
ci.plot(simZ_genY_white, ylim = 0:1, ylab = "Probability of voting", 
        main = "White Millennial Man")
#Now, Baby Boomer
setZ_boomer_white <- setx(Z_boomer, income = 1:7, education = 2, black = 0, 
                          hispanic = 0, female =0, vote_2008 = 1)
simZ_boomer_white <- sim(Z_boomer, x = setZ_boomer_white)
ci.plot(simZ_boomer_white, ylim = 0:1, ylab = "Probability of voting", 
        main = "White Baby Boomer Man")
```

###  white, high-school-educated man across levels of income, assuming he didn't vote in 2008. 

```{r, message=FALSE, echo = FALSE, fig.align='center', fig.show='hold', fig.width=3.5, fig.height=5}
setZ_genY_white <- setx(Z_genY, income = 1:7, education = 2, black = 0, 
                        hispanic = 0, female =0, vote_2008 = 1)
simZ_genY_white <- sim(Z_genY, x = setZ_genY_white)
ci.plot(simZ_genY_white, ylim = 0:1, ylab = "Probability of voting", 
        main = "White Millennial Man")
#Now, Baby Boomer
setZ_boomer_white <- setx(Z_boomer, income = 1:7, education = 2, black = 0, 
                          hispanic = 0, female =0, vote_2008 = 1)
simZ_boomer_white <- sim(Z_boomer, x = setZ_boomer_white)
ci.plot(simZ_boomer_white, ylim = 0:1, ylab = "Probability of voting", 
        main = "White Baby Boomer Man")
```

### Plot 

```{r, message=FALSE, echo = FALSE, fig.align='center', fig.show='hold', fig.width=3.5, fig.height=5}
ci.plot(simZ_genY_white, ylim = 0:1, ylab = "Probability of voting", 
        main = "White Millennial Man")
Z_genX <- zelig(vote_2012 ~ education + female + vote_2008 + black + hispanic + income, model = "logit.survey", weights=~weights, strata=~strata, 
                  data = anes_genX)

setZ_genX_white <- setx(Z_genX, income = 1:7, education = 2, black = 0, 
                        hispanic = 0, female =0, vote_2008 = 0)
simZ_genX_white <- sim(Z_genX, x = setZ_genX_white)
ci.plot(simZ_genY_white, ylim = 0:1, ylab = "Probability of voting", 
        main = "White Gen X Man")
```

### plot

```{r, message=FALSE, echo = FALSE, fig.align='center', fig.show='hold', fig.width=3.5, fig.height=5}
setZ_genY_voted <- setx(Z_genY, income = 1:7, education = 2, black = 0, 
                        hispanic = 0, female =0, vote_2008 = 1)
simZ_genY_voted <- sim(Z_genY, x = setZ_genY_voted)
ci.plot(simZ_genY_voted, ylim = 0:1, ylab = "Probability of voting", 
        main = "White Millennial man who voted in '08")
setZ_genX_voted <- setx(Z_genX, income = 1:7, education = 2, black = 0, 
                        hispanic = 0, female =0, vote_2008 = 1)
simZ_genX_voted <- sim(Z_genX, x = setZ_genX_voted)
ci.plot(simZ_genX_voted, ylim = 0:1, ylab = "Probability of voting", 
        main = "White Gen X man who voted in '08")
```
### The effects of differences in voting history only increase with age. 
```{r, message=FALSE, echo = FALSE, fig.align='center', fig.show='hold', fig.width=3.5, fig.height=5}
ci.plot(simZ_genX_voted, ylim = 0:1, ylab = "Probability of voting", 
        main = "White Gen X man who voted in '08")
setZ_boomer_voted <- setx(Z_boomer, income = 1:7, education = 2, black = 0, 
                          hispanic = 0, female =0, vote_2008 = 1)
simZ_boomer_voted <- sim(Z_boomer, x = setZ_boomer_voted)
ci.plot(simZ_boomer_voted, ylim = 0:1, ylab = "Probability of voting", 
        main = "White Baby Boomer man who voted in '08")
```
### The effects of differences in voting history only increase with age. 
```{r, message=FALSE, echo = FALSE, fig.align='center', fig.show='hold', fig.width=3.5, fig.height=5}
ci.plot(simZ_boomer_voted, ylim = 0:1, ylab = "Probability of voting", 
        main = "White Baby Boomer man who voted in '08")

setZ_silent_voted <- setx(Z_silent, income = 1:7, education = 2, black = 0, 
                          hispanic = 0, female =0, vote_2008 = 1)
simZ_silent_voted <- sim(Z_silent, x = setZ_silent_voted)
ci.plot(simZ_silent_voted, ylim = 0:1, ylab = "Probability of voting", 
        main = "White Silent Generation man who voted in '08")
```


## Model results {.flexbox .vcenter .smaller} 
 
-Expressed as odds-ratios (exponentiated coefficients), so
can be interpreted as factor change in odds of voting in 2012
```{r echo = FALSE, results='asis', message=FALSE,}
#setwd('C:/Users/Chris/Documents/GitHub/Final_Presentation')
library(stargazer)
stargazer(M_genY, M_genX, M_boomer, M_silent, apply.coef = exp,
         t.auto=F, p.auto=F, report = "vct*", font.size = 'tiny',
          type = "html",  dep.var.labels = "Probability of voting in 2012", 
          column.labels = c("Gen Y", "Gen X", "Boomers", "Silent"),
          title = "Odds Ratios Across Four Generations")
```



     
## Limitations 

>- Issues with external validity
>- Relavtively low survey response 
>- Self survey & self report
>- Goodness of Fit  

## Goodness of Fit (Likelihood-ratio test)
Does the model with predictors fit significantly better than a model with just an intercept (a null model)? 
```{r, echo=TRUE}
with(M_genY, null.deviance - deviance) #gives chi-square value
with(M_genY, df.null - df.residual) #gives degrees of freedom
with(M_genY, pchisq(null.deviance - deviance, 
                    df.null - df.residual, 
                    lower.tail = FALSE)) #gives p value
```
## Goodness of Fit (ROC)
Receiver operating characteristic plot
```{r}

rocplot(Z_genY, Z_boomer, col1 = 'blue', col2 = 'red')
legend("left", c("Millennial","Baby Boomer"), 
       lty=c(1,1), lwd=c(2.5,2.5), col=c('blue','red'))

```
